---
title: "exploreCHVIs"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

```

## Preparing the Data
This section installs the necessary pacakage(s) and then loads a function ('readCHVI') which will 
1: read in an individual CHVI csv
2: limit the columns to only those (26) of interest
3: reorder those columns alphabetically so that they can be renamed identically
4: rename the columns
4a: clean up the report year field so it is handled correctly

```{r prepare}
rm(list=ls())
library(tidyverse)


readCHVI <- function(fileName){
   temp <- read.csv(fileName, header=T, stringsAsFactors = F) %>% select(1:26) %>% .[,order(names(.))]
   
   names(temp) <- c(
     "CA_decile",
     "CA_RR",
     "county_fips",
     "county_name",
     "denominator",
     "estimate",
     "geoname",
     "geotype",
     "geotypevalue",
     "ind_definition",
     "ind_id",
     "LL_95CI",
     "numerator",
     "race_eth_code",
     "race_eth_name",
     "region_code",
     "region_name",
     "reportyear",
     "RSE",
     "SE",
     "strata_one_code",
     "strata_one_name",
     "strata_two_code",
     "strata_two_name",
     "UL_95CI",
     "version"
   )
   
   temp$reportyear <- as.character(as.numeric(temp$reportyear))
   
   return(temp)
}

```

## Read in individual CHIV csvs

This portion reads in each of the CHVI csv files and then combines them all into a single dataframe. Set the working directory to fit your workflow.


```{r readem}

# work Local
setwd("~/CHVI_copy/CHVIcsvs/")

# work from the network
#setwd("//phitprlcsrvip04/OHEGroup/HCI/Data/CCHVI in one folder for web/CHVIcsvs/")

airConditioning <- readCHVI("./BRACE_AirConditioning_797_CO_RE_CA.csv")
carOwnership <- readCHVI("./BRACE_CarOwnership_37_CT_PL_CO_RE_CA.csv")
children <- readCHVI("./BRACE_children_788_CT_PL_CO_RE_CA.csv")
disability <- readCHVI("./BRACE_Disability_795_CT_PL_CO_RE_CA.csv")
elderly <- readCHVI("./BRACE_elderly65over_789_CT_PL_CO_RE_CA_11-8-2016.csv")
extremeHeat <- readCHVI("./BRACE_ExtremeHeat_791_CO.csv")
impervious <- readCHVI("./BRACE_ImperviousSurfaces_423_CT_PL_CO_RE_CA.csv")
insurance <- readCHVI("./BRACE_Insurance_795_CT_PL_CO_RE_CA.csv")
linguisticIso <- readCHVI("./BRACE_LinguisticIsolation_800_CT_PL_CO_RE_CA.csv")
outdoorWorkers <- readCHVI("./BRACE_OutdoorWorkers_790_CT_PL_CO_RE_CA.csv")
ozone <- readCHVI("./BRACE_Ozone_801_CT_PL_CO_RE_CA.csv")
pm25 <- readCHVI("./BRACE_PM25levels_776_CT_PL_CO_RE_CA.csv")
race <- readCHVI("./BRACE_race_795_CT_PL_CO_RE_CA.csv")
seaLevel <- readCHVI("./BRACE_SLR_784_CT_PL_CO_RE_CA_11-1-2016.csv")
treeCanopy <- readCHVI("./BRACE_TreeCanopy_458_CT_PL_CO_RE_CA.csv")
wildfire <- readCHVI("./BRACE_Wildfire_786_CT_PL_CO_RE_CA.csv")


allCHVI <- bind_rows(airConditioning,
                     carOwnership, 
                     children, 
                     disability, 
                     elderly, 
                     extremeHeat, 
                     impervious, 
                     insurance, 
                     linguisticIso,
                     outdoorWorkers, 
                     ozone, 
                     pm25, 
                     race,
                     seaLevel,
                     treeCanopy, 
                     wildfire
                     ) %>%
  filter(ind_definition != "")

head(allCHVI)

```

## Produce a sample plot

You can now query the data to obtain the data (ie indicator, county, strata, etc.) of interest and plot the results. 

```{r samplePlot1}
allCHVI %>% 
  filter(geotype == "CA" & estimate > 0 & ind_definition == "Percent of households with air conditioning" & race_eth_name !="Total")  %>% 
  ggplot(aes(x=county_name, y=estimate, fill=race_eth_name)) + geom_bar(stat = "identity", position="dodge") + ggtitle("CA % of households with air conditioning (by race)")  
  
```


## Add error bars 

Here we add error bars to the plot previously produced. 

```{r samplePlot2}

allCHVI %>% 
  filter(geotype == "CA" & estimate > 0 & ind_definition == "Percent of households with air conditioning" & race_eth_name !="Total") %>%
  ggplot(aes(x=county_name, y=estimate,ymax = UL_95CI, ymin=LL_95CI, fill=race_eth_name)) + geom_bar(stat = "identity", position="dodge") + ggtitle("CA % of households with air conditioning (by race)") + geom_errorbar(position="dodge")
```

## Examine the data

Something looks weird. The result is concerning because the 'estimates' do no always fall within the confidence intervals...

We can verify this by looking at the data that we send to the plot 

```{r examine1}
allCHVI %>% 
  filter(geotype == "CA" & estimate > 0 & ind_definition == "Percent of households with air conditioning" & race_eth_name !="Total") %>%
  select(LL_95CI, estimate, UL_95CI)
  
```


## Reproduce a graph from the VA report

This will grab data on extreme heat days and produce a new chart. 

```{r examineEH}
# get list of all available indicators
unique(allCHVI$ind_definition)

# then examine the records
allCHVI %>% 
  filter(geotype == "CO" & ind_definition == "Projected number of extreme heat days") %>% head()
```
It looks like we have two records (report years) for each county, and we can use 'estimate' to get the values we are interested in seeing. 
```{r sampleEHplot, out.width = '\\maxwidth'}
allCHVI %>% 
  filter(geotype == "CO" & ind_definition == "Projected number of extreme heat days") %>% 
  ggplot(aes(x=county_name, y=estimate, fill=region_name)) + geom_bar(stat = "identity", position="dodge") + ggtitle("# of projected Extreme Heat Days")  + facet_grid(.~ reportyear) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.25, size=5))


```

Or look at just the Bay Area and Southern Califronia counties to make things less crowded. 
```{r sampleEHplot3, out.width = '\\maxwidth'}
allCHVI %>% 
  filter(geotype == "CO" & ind_definition == "Projected number of extreme heat days" &  region_name %in% c("Bay Area","Southern California")) %>% 
  ggplot(aes(x=county_name, y=estimate, fill=reportyear)) + geom_bar(stat = "identity", position="dodge") + ggtitle("# of projected Extreme Heat Days")  + facet_grid(.~ region_name) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.25, size=8))


```